---
title: "Activity monitoring data"
author: "Jingtao Lian"
date: "15-Jan-2017"
---
# Reproducible Research: Peer Assessment 1

## 1. Loading and preprocessing the data
```{r}
library("ggplot2")  
act <- read.csv("activity.csv")
```

## 2. Histogram of the total number of steps taken each day
```{r}
agg <- aggregate(steps~date, act, sum)
ggplot(agg, aes(steps)) + geom_histogram(binwidth = 1000)
```

## 3. Mean and median number of steps taken each day
```{r}
mean(agg$steps)
median(agg$steps)
```

## 4. Time series plot of the average number of steps taken
```{r}
step_interval <- aggregate(steps ~ interval, act, mean)
ggplot(step_interval, aes(interval, steps)) + geom_line() + xlab("5-minute interval") + 
    ylab("average number of steps taken")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
step_interval[which.max(step_interval$steps), ]
```

## 6. Code to describe and show a strategy for imputing missing data
```{r}
missing <- is.na(act$steps)
table(missing)
```
## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
imputed <- function (step, interval) {
    filled <- NULL
    if (!is.na(step)) {
        filled <- step
    } else {
        filled <- mean(step_interval$steps[step_interval$interval == interval])
    }
    return(filled)
}

filledDate <- act
filledDate$steps <- mapply(imputed, filledDate$steps, filledDate$interval)
stepsByDate <- aggregate(steps~date, filledDate, sum)
hist(stepsByDate$steps, xlab = "steps", ylab = "counts", main = "total number of steps taken each day")

```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
weekdaysOrWeekend <- function(day) {
    day <- weekdays(day)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) {
        return("weekday")
    } else {
        return("weekend")
    }
}

filledDate$date <- as.Date(filledDate$date)
filledDate$date <- mapply(weekdaysOrWeekend, filledDate$date)

aggByDate <- aggregate(steps~interval + date, filledDate, sum)
ggplot(aggByDate, aes(interval, steps)) + geom_line() + facet_grid(date ~ .) + 
    xlab("5 minutes interval") + ylab("Numbers of Steps")
```